---
title: "Codebook"
author: "Jayant"
date: "1 July 2016"
output: html_document
---
First set the working directory

```{r}
setwd("~/Coursera/Data Science/Getting&Cleaning_Data/Project")
```

Create a dir called "data" if it does not exist Create it
Download the file and put it in the "data" directory

```{r}
if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileUrl,destfile="./data/Dataset.zip")
```
Unzip the File

```{r}
unzip(zipfile="./data/Dataset.zip", exdir="./data")
```
unzipped files are in the folder "UCI HAR Dataset"

Get the list of the files
```{r}
path_rf <- file.path("./data" , "UCI HAR Dataset")
files<-list.files(path_rf, recursive=TRUE)
files
```

Read data from the files into the variables

Read the Activity files

```{r}
dataActivityTest  <- read.table(file.path(path_rf, "test" , "Y_test.txt" ),header = FALSE)
dataActivityTrain <- read.table(file.path(path_rf, "train", "Y_train.txt"),header = FALSE)
```

Read the Subject files

```{r}
dataSubjectTrain <- read.table(file.path(path_rf, "train", "subject_train.txt"),header = FALSE)
dataSubjectTest  <- read.table(file.path(path_rf, "test" , "subject_test.txt"),header = FALSE)
```

Read Fearures files

```{r}
dataFeaturesTest  <- read.table(file.path(path_rf, "test" , "X_test.txt" ),header = FALSE)
dataFeaturesTrain <- read.table(file.path(path_rf, "train", "X_train.txt"),header = FALSE)
```

Look at the varibles

```{r}
str(dataActivityTest)
```

```{r}
str(dataActivityTrain)
```

```{r}
str(dataSubjectTrain)
```

```{r}
str(dataSubjectTest)
```


```{r}
str(dataFeaturesTest)
```

```{r}
str(dataFeaturesTrain)
```


************** Assignment 1 ***************

Merges the training and the test sets to create one data set

1.Concatenate the data tables by rows

```{r}
dataSubject <- rbind(dataSubjectTrain, dataSubjectTest)
dataActivity<- rbind(dataActivityTrain, dataActivityTest)
dataFeatures<- rbind(dataFeaturesTrain, dataFeaturesTest)
```

2.set names to variables

```{r}
names(dataSubject)<-c("subject")
names(dataActivity)<- c("activity")
dataFeaturesNames <- read.table(file.path(path_rf, "features.txt"),head=FALSE)
names(dataFeatures)<- dataFeaturesNames$V2
```

3. Merge columns to get the data frame "MasterData" for all data

```{r}
dataCombine <- cbind(dataSubject, dataActivity)
MasterData <- cbind(dataFeatures, dataCombine)
```

************** Assignment 2 ***************
Extracts only the measurements on the mean and standard deviation for each measurement
Subset Name of Features by measurements on the mean and standard deviation
i.e taken Names of Features with "mean()" or "std()"

```{r}
NamesFeatSubdata<-dataFeaturesNames$V2[grep("mean\\(\\)|std\\(\\)", dataFeaturesNames$V2)]
```

Subset the data frame "MasterData" by seleted names of Features
```{r}
selectedNames<-c(as.character(NamesFeatSubdata), "subject", "activity" )
MasterData<-subset(MasterData,select=selectedNames)
```

See the structure of the dataframe

```{r}
str(MasterData)
```

************** Assignment 3 ***************
Uses descriptive activity names to name the activities in the data set

```{r}
MasterData$activity <- as.character(MasterData$activity)
MasterData$activity[MasterData$activity == 1] <- "WALKING"
MasterData$activity[MasterData$activity == 2] <- "WALKING_UPSTAIRS"
MasterData$activity[MasterData$activity == 3] <- "WALKING_DOWNSTAIRS"
MasterData$activity[MasterData$activity == 4] <- "SITTING"
MasterData$activity[MasterData$activity == 5] <- "STANDING"
MasterData$activity[MasterData$activity == 6] <- "LAYING"
MasterData$activity <- as.factor(MasterData$activity)
```

Check

```{r}
head(MasterData$activity, 30)
```


************** Assignment 4 ***************
Appropriately labels the data set with descriptive variable names.
In the former part, variables activity and subject and names of the activities 
have been labelled using descriptive names.In this part, Names of Feteatures will 
labelled using descriptive variable names.

<ul>
<li>  prefix t is replaced by time</li> 
<li>  Acc is replaced by Accelerometer </li>
<li>  Gyro is replaced by Gyroscope </li>
<li>  prefix f is replaced by frequency </li> 
<li>  Mag is replaced by Magnitude </li>
<li>  BodyBody is replaced by Body </li>
</ul>

```{r}
names(MasterData)<-gsub("^t", "time", names(MasterData))
names(MasterData)<-gsub("^f", "frequency", names(MasterData))
names(MasterData)<-gsub("Acc", "Accelerometer", names(MasterData))
names(MasterData)<-gsub("Gyro", "Gyroscope", names(MasterData))
names(MasterData)<-gsub("Mag", "Magnitude", names(MasterData))
names(MasterData)<-gsub("BodyBody", "Body", names(MasterData))
```

Check

```{r}
names(MasterData)
```


************** Assignment 5 ***************
From the data set in step 4, creates a second, independent tidy data set with the average 
of each variable for each activity and each subject.

Use the required Library

```{r, message=FALSE, warning=FALSE}
library(plyr);
library(data.table);
library(dplyr);
```

```{r}
MasterDAta2 <- data.table(MasterData)
```

This takes the mean of every column broken down by participants and activities

```{r}
TidyData <- MasterDAta2[, lapply(.SD, mean), by = 'subject,activity']
write.table(TidyData, file = "Tidy_Data.txt", row.names = FALSE)
```

Done!
